---
title: " DATA1 : CLINICAL DATA/gen ON AD SUBJETCS"
author: "GAB"
date: "10/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
MAARS_AD_full_20190131_12-34-49.csv  
= Clinical and technical features specific to AD samples (only AD Lesional and AD Non_lesional)  
Please be careful that clinical and technical features are common for a same MAARS_identifier and sample_id.  
Thus, for the severity stratification (SCORAD_Score) it is recommended to analyse separately lesional and non lesional skin.
Most interested variables for sample identification:  
● involved skin biopsy#involved skin biopsy#MAARS Sample identifier (MAARS_Sample_identifier) : lesional sample_id    
● uninvolved skin biopsy#uninvolved skin biopsy#MAARS Sample identifier (MAARS_Sample_identifier) : non lesional sample_id    
Most interesting variable for severity stratification and prediction  
● patient#SCORAD index#SCORAD#SCORAD Score (SCORAD_Score)  
The others variables can be useful to identify potential confounding factor of interesting clinical features  

## importing data
```{r importing AD dataset}
library(tidyverse)
library(readxl)
library(knitr)
data1 <- read_excel("M2PHDS_19-20_OMICS_CLIN_DATA_MAARS_AD_full_20190131_12-34-49_backup.xlsx")
```
94 patients included with AD but 91 were included according to the paper.

## Subsetting
### Patient identification subset: A1 to A4
```{r subsetting data, identification}
patient_identification_df <- data1[ , 1:4]
patient_identification_df %>%
   group_by(A4) %>%
      count(patient_id= paste(A1, collapse = ","))
```
39 patients are from UHU: Dusseldorf Germany
16 from KINGS, London
39 from UH, Helsinki, Finland  

### Inclusion/exclusion criteria subset: A5 to A13
```{r subsetting data, inclusion/exclusion criteria}
inclusion_df <- data1[ , c(1, 5:13)]
```


```{r checking A11}
inclusion_df %>% 
  group_by(A11)%>%
  count(A1)
inclusion_df_clean <- inclusion_df %>%
filter(A11=="No")#
```
One patient had antibiotics during the previous 2 weeks: 2 weeks MAARS_2_013, removing him

check inclusion_df_clean: MAARS_2_013 is gone
```{r checking A5 Written_informed_consent}
inclusion_df_clean %>%
  group_by(A5) %>%
  count(A1)

```
All of participants gave written informed consent  

```{r checking A6, 18_yo}
inclusion_df_clean %>%
  group_by(A6) %>%
  count(A1)
```
All participants were 18 or more 


```{r checking A7, Patients_with_Atopic_Dermatitis,_diagnosed_using_Hanifin-Rajka_Criteria}
inclusion_df_clean %>%
  group_by(A7) %>%
  count(A1)
```
All participants were diagnosed with SCORAD

```{r checking A8, Subject_unable_to_give_written_informed_consent}
inclusion_df_clean %>%
  group_by(A8) %>%
  count(A1)
```
All participants were able to give written informed consent

```{r checking A9, Subject_has_no_allergen-specific_IgE_and_no_allergen-specific_immediate_type_reactions}
inclusion_df_clean %>%
  group_by(A9) %>%
  count(A1)
```
All participants didn't have allergen-specific IgE

```{r checking A10, Subjects_who_have_received_treatment_at_the_biopsy_site_at_least_2_weeks_prior_to_screening}
inclusion_df_clean %>%
  group_by(A10) %>%
  count(A1)
```
All participants didn't receive treatment at the biopsy site at least 2 weeks prior to screening 

*One subject removed based on inclusion/exclusion criteria*

### Diagnostic & Phenotypic Data: A14 to A20

```{r subsetting Diagnostic and phenotypic data}
diag_pheno_df <- data1[ , c(1, 14:20)]
```

```{r A14, looking at subjects date of birth}
kable(diag_pheno_df %>%
  summarise(mean(A14), sd(A14)))
```

```{r A15, ethnicity}
kable(diag_pheno_df %>%
  count(A15))
```

### Hanifin and Rajka diagnostic criteria: A21 to A57
```{r subsetting Hanifin and Rajka diagnostic criteria}
scorad_criteria <- data1[ , c(1, 21:57)]
```

### SCORAD index and scores: A215 to A227
```{r subsetting SCORAD index and scores}
scorad_score <- data1[ , c(1, 215:227)]

```

```{r SCORAD total scores A226}
kable(scorad_score$A226)

kable(scorad_score %>%
  summarise(mean(A226), sd(A226)))
```

### Subsetting for Other concurrent chronic diseases: A189 to 208
```{r Subsetting for Other concurrent chronic diseases}
chronic_disease <- data1[ , c(1, 189:208)]
```

### Subsetting for exlcuded categories: A252, 264,269, 276, 284, 288, 292, 296

```{r Subsetting for exlcuded categories}
excluded <-dplyr::select(data1, A1, A252, A260, A264, A268, A272, A276, A280,A284, A288, A292, A296)
```


```{r excluded from curated items: A252}
kable(excluded%>%
        group_by(A1)%>%
        count(A252=="Yes"))
```
5 patients excluded from curated items

```{r excluded from involved skin biopsy: A264}
kable(excluded%>%
  count(A264=="Yes"))
```
all patients kept from involved skin biopsy

```{r excluded from uninvolved skin biopsy: A268}
kable(excluded%>%
        group_by(patient_id)%>%
        
  count(A268=="Yes"))
```
1 patient excluded from uninvolved skin biopsy

```{r excluding patients}
data1 <- data1 %>%
  filter(A11 == "No", is.na(A252), is.na(A268))

```

88 patients remaining
descriptive analysis


```{r}
```

