---
title: "data management for file all AD"
author: "GAB"
date: "10/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
library("readxl")
data2 <- read_excel("M2PHDS_19-20_OMICS_CLIN_DATA_MAARS_all_Fri_Apr_04_14h_CEST_2014.xlsx")
head(data2)
```

Selecting subjects and removing redundancy: one single identifier 
```{r grouping by subject/ MAARS_identifier}
data2_grouped <-data2 %>%
  distinct(MAARS_identifier, .keep_all=TRUE) %>%
  arrange(MAARS_identifier)

head(data2_grouped)
data2_grouped %>%
  count(clinical_group)
```
That leaves us with 334 people instead of 316 as initially in the flow chart
88 AD like in the full AD file after excluding patients. 

```{r}
data2_AD <- filter(data2_grouped, clinical_group=="AD")
intersect(data2_grouped$MAARS_identifier, data1$A1)
setdiff(data2_AD$MAARS_identifier,data1$A1)

```
91 AD patient in data2

that leaves 87 participants in common 
"MAARS_2_104" is not in common between the 2 databases, is in data_full_clean but not in data_all_AD
removing him from data_all:
```{r}
data1<- data1 %>%
  filter(A1 != "MAARS_2_104")
```

```{r}
setdiff(data2_AD$MAARS_identifier,data1$A1)
data2_AD <- data2_AD %>%
  filter(MAARS_identifier != "MAARS_1_017")
```
"MAARS_1_017 is in data_all_AD but not in data_full_clean anymore, 
we need to remove him from the data2_AD



