---
title: "OMICS Project"
author: "Arulmani, Gabriel, Penn"
date: "10/13/2019"
output: 
  html_document:
    toc: true
    number_sections: true  
    
fontsize: 12pt 

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installing the Geo query bio conductor package
```{r, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

```


# Importing the libraries required
```{r, eval=FALSE}
BiocManager::install("GEOquery")
BiocManager::install("edgeR")
BiocManager::install("limma")

install.packages("UpSetR")
install.packages("flashClust")
library(readr)
library(dplyr)
library(limma)
library(edgeR)
library(UpSetR)
library(flashClust)
library(readxl)
library(GEOquery)
```



# Importing the dataset

```{r}

AD_all_data <- read.csv(file = "M2PHDS_19-20_OMICS_CLIN_DATA_MAARS_all_Fri_Apr_04_14h_CEST_2014.csv", header = TRUE, sep = "\t")
MAARS_all_data <- read.csv(file = "M2PHDS_19-20_OMICS_CLIN_DATA_MAARS_AD_full_20190131_12-34-49.csv", header = TRUE, sep = "\t")
omics_data <- read.delim("/home/user-02/Desktop/Project_OMICS/M2PHDS_19-20_OMICS_TRANSC_MAARS_normTranscriptome_618samples_16042014.txt")

```


## Checking the dimensions of the given dataset and variables names
```{r}
dim(AD_all_data)
str(AD_all_data, eval = FALSE)
dim(MAARS_all_data)
str(MAARS_all_data, eval = FALSE)
dim(omics_data)
str(omics_data, eval = FALSE)
```

```{r, eval=FALSE}
names(AD_all_data)
names(MAARS_all_data)
```

## Calculating samples in each sub group
```{r}
library(dplyr)
AD_all_data %>% group_by(AD_all_data$clinical_group)%>% summarise(n = n())

# Calculating the distinct individual in each group
AD_all_data %>% group_by(AD_all_data$clinical_group)%>% summarise(n = n_distinct(MAARS_identifier))

```

## Calculating the samples as per the lesion status

```{r}
AD_all_data%>%group_by(AD_all_data$lesional) %>% summarise(n = n())

#calculating the distinct samples in each group
AD_all_data%>% group_by(AD_all_data$lesional)%>% summarise(n = n_distinct(MAARS_identifier))

#Calculating the distinct samples in each institution
AD_all_data%>%group_by(AD_all_data$Institution)%>%summarise(n_distinct(MAARS_identifier))
# Calculating the distinct samples under anatomical location
AD_all_data%>% group_by(AD_all_data$anatomical_location_label)%>%summarise(n_distinct = n_distinct(MAARS_identifier))
```

# Analysing the other clinical covariates descriptives
```{r}

Descriptives_AD_data<- AD_all_data %>% group_by(AD_all_data$clinical_group)%>% 
  summarise( n_distinct = n_distinct(MAARS_identifier), percentage_anatomical_loc_PT = sum(anatomical_location_label == "PT")/n()*100, percentage_anatomical_loc_UB = sum(anatomical_location_label == "UB")/n()*100, percentage_anatomical_loc_B = sum(anatomical_location_label == "B")/n()*100, percentage_anatomical_loc_LB = sum(anatomical_location_label == "LB")/n()*100, percentage_anatomical_loc_T = sum(anatomical_location_label == "T")/n()*100, percentage_HHU = sum(Institution == "HHU")/n()*100, percentage_KINGS = sum(Institution == "KINGS")/n()*100, percentage_UH = sum(Institution == "UH")/n()*100, mean_age = mean(CUSTOM_Age), sd_age = sd(CUSTOM_Age), percentage_male  = sum(Gender == "Male")/n()*100)
Descriptives_AD_data

```

# Checking the allergies variables
```{r}


Descrptives_AD_data_allergy<- AD_all_data%>% group_by(AD_all_data$clinical_group)%>% 
  summarise( n_distinct = n_distinct(MAARS_identifier), percentage_pseudo_allergy = sum(Known_Allergies_v2..Pseudo_Drug_Allergy == "FALSE")/n()*100, percentage_House_dust_mite = sum(Known_Allergies_v2..House_dust_mite)/n()*100,  percentage_Food = sum(Known_Allergies_v2..Food)/n()*100,  percentage_Pollen = sum(Known_Allergies_v2..Pollen)/n()*100,  percentage_contact_allergy = sum(Known_Allergies_v2..Contact_Allergy)/n()*100,  percentage_drug_allergy = sum(Known_Allergies_v2..Drug_Allergy)/n()*100,  percentage_allergy_animal = sum(Known_Allergies_v2..Animal)/n()*100)
Descrptives_AD_data_allergy
```


# Checking the differences among groups(AD, PSO, CTRL)
### Barplot of the library sizes of each sample
```{r}
clinical_group <- AD_all_data$clinical_group
levels(clinical_group) <- c("blue2", "gold", "green2")
boxplot(log2(omics_data),
    col = as.character(clinical_group), xlab = "Samples", ylab = "log2(raw counts)", 
    axes = FALSE)
axis
box()
legend("topright", c("Atopic Dermatits", "P(2)soriasis", "Healthy Volunteers"), col = c("blue2", "gold", "green2"), pch = 15, horiz = TRUE, bg= NA)

```


## Checking the differences in lesion status
```{r}
lesion_status <- AD_all_data$lesional
levels(lesion_status) <- c("blue2", "gold")
boxplot(log2(select(omics_data, -1)),
        col = as.character(lesion_status), xlab = "Samples", ylab = "log2(raw counts)",
        axes = FALSE)
axis(2)
box()
legend("topright", c("Lesional", "Nonlesional"), col = c("blue2", "gold"), pch = 15, horiz = TRUE, bg = NA)
```

It is inferred that the plot shows us the data is reasonably normalized with no library size effects


## Checking the sequecning depth diffrences in clinical group
```{r}
barplot(colSums(select(omics_data, -1)), 
    col = as.character(clinical_group), xlab = "Samples", ylab = "Library size", names.arg = "")
legend("topright", c("Atopic Dermatits", "Psoriasis", "Healthy Volunteers"), col = c("blue3", 
    "gold", "green2"), pch = 15, horiz = FALSE, bg = NA)
```

## Checking the sequecning depth diffrences in lesion status

```{r}
barplot(colSums(select(omics_data, -1)), 
    col = as.character(lesion_status), xlab = "Samples", ylab = "Library size", names.arg = "")
legend("topright", c("Lesional", "Nonlesional"), col = c("blue3", 
    "gold"), pch = 15, horiz = FALSE, bg = NA)
```
It is inferred that the there is no sequencing depth problem which could have been treated with TMM normalization

## Dendrogram of the hierarchial clusterings using ward method and euclidean distance

```{r, eval=FALSE}
deucl_norm <- dist(t(omics_data))
plot(deucl_norm)

colnames(omics_data)[1:10]
rownames(omics_data)[1:15]

plot(flashClust::hclust(deucl_norm, method = "ward"), labels = AD_all_data$clinical_group)

```

## Adding the omics modified file 
```{r, eval= FALSE}

omics_modified_data<- read.delim("~/Desktop/Project_OMICS/M2PHDS_19-20_OMICS_TRANSC_MAARS_normTranscriptome_618samples_16042014.csv")
gsub(pattern = )
```


